---
title: "STA 445 HW2"
author: "Matthew Hashim"
date: "2/27/2004"
output: pdf_document
---

## Problem 1

Create a vector of three elements (2,4,6) and name that vector `vec_a`. Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors together and name the result `vec_c`.

```{r}
vec_a = c(2, 4, 6)
vec_b = c(8, 10, 12)
vec_c = vec_a + vec_b
vec_c
```

## Problem 2

Create a vector, named `vec_d`, that contains only two elements (14,20). Add this vector to `vec_a`. What is the result and what do you think R did (look up the recycling rule using Google)? What is the warning message that R gives you?

```{r}
vec_d = c(14, 20)
vec_a + vec_d
```
The warning was "longer object length is not a multiple of shorter object length".
I think that r will continue to add vec_d to vec_a until every single index in a has had a value added. So index 0 and 1 got added by vec_d, then 2 and 3, until the end of vect_a. But since vec_a was length 3 we could never add a full multiple of vect_d.

## Problem 3

Next add 5 to the vector vec_a. What is the result and what did R do? Why doesn't in give you a warning message similar to what you saw in the previous problem?

```{r}
vec_a + 5
```

It consideres 5 a vector of length 1. And any 1 length vector is going to be a multiple of any other vector.

## Problem 4

Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two different ways. 

a.  First using the `seq()` function

```{r}
seq(1, 5)
```

b. Using the `a:b` shortcut.

```{r}
1:5
```

## Problem 5

Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 

a. Using the seq() function

```{r}
seq(2, 6, 2)
```

b. Using the a:b shortcut and some subsequent algebra. 

```{r}
1:3 * 2
```

## Problem 6

Generate a vector of 21 elements that are evenly placed between 0 and 1 using the `seq()` command and name this vector `x`. 

```{r}
x = seq(0, 1, length.out = 21)
x
```


## Problem 7

Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ using the `rep()` command to replicate the vector c(2,4,8). 
  
```{r}
rep(c(2, 4, 8), 3)
```

## Problem 8
Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
using the `rep()` command. You might need to check the help file for rep() to see all of the options that rep() will accept. In particular, look at the optional argument `each=`.
  
```{r}
rep(c(2, 4, 8), each = 4)
```
  
## Problem 9 

In this problem, we will work with the matrix 
$$ \left[\begin{array}{ccccc}
    2 & 4 & 6 & 8 & 10\\
    12 & 14 & 16 & 18 & 20\\
    22 & 24 & 26 & 28 & 30
    \end{array}\right]$$
    
a.  Create the matrix in two ways and save the resulting matrix as `M`.

i. Create the matrix using some combination of the `seq()` and `matrix()` commands.

```{r}
M = matrix(seq(2, 30, 2), 3, 5, byrow = TRUE)
M
```

ii. Create the same matrix by some combination of multiple `seq()` commands and either the `rbind()` or `cbind()` command.

```{r}
M = rbind(seq(2, 10 , 2), seq(12, 20, 2), seq(22, 30, 2))
M
```

b. Extract the second row out of M.

```{r}
M[2,]
```

c. Extract the element in the third row and second column of M

```{r}
M[3,2]
```

## Problem 10 

The following code creates a `data.frame` and then has two different methods for removing the rows with `NA` values in the column `Grade`. Explain the difference between the two.

```{r, eval=FALSE}
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
df[ -which(  is.na(df$Grade) ), ]

df[  which( !is.na(df$Grade) ), ]
```

The first method has a - before which. This means that it will only keep those that are not true.
The second method has a not symbol before is.na. This means that it will make those that are NA into false and only the true values are returned.

## Problem 11 

Create and manipulate a list.

a.  Create a list named my.test with elements
        + x = c(4,5,6,7,8,9,10)
        + y = c(34,35,41,40,45,47,51)
        + slope = 2.82
        + p.value = 0.000131
```{r}
my.test = list(x = c(4,5,6,7,8,9,10), y = c(34,35,41,40,45,47,51),
  slope = 2.82, p.value = 0.000131)
```

b.  Extract the second element in the list.

```{r}
my.test[2]
```

c.  Extract the element named `p.value` from the list.

```{r}
my.test$p.value
```



